Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Clustering and Parameter Estimation

### 2020-09-30 summary.Rmd
temporary push

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}

@@

*Proof*. 
Let $\Lambda^{(k)} = 
\begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
\in \mathbb{R}^{n_k \times K}$.  
Let $X$ be a block diagonal matrix 
$X = \diag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}$.  
Let $L^{(k)}$ be a block diagonal matrix of column vectors $\lambda^{(lk)}$ for 
$l = 1, ..., K$. $L^{(k)} = \diag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
\mathbb{R}^{n \times K}$.  
Let $Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
\mathbb{R}^{n \times K^2}$.  
Then $P = X Y^\top$.  
Similar to the $K = 2$ case, we again have $Y = X \Pi$ for a permutation matrix
$\Pi$, so $P = X \Pi X^\top$.  
The permutation described by $\Pi$ has $K$ fixed points, which correspond to 
$K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where 
$k = r (K + 1) + 1$ for $r = 0, ..., K - 1$. It also has 
$\binom{K}{2} = K (K - 1) / 2$ cycles of order $2$. Each cycle corresponds to 
a pair of eigenvalues $+1$ and $-1$ and a pair of eigenvectors 
$(e_s + e_t) / \sqrt{2}$ and $(e_s - e_t) / \sqrt{2}$.

@@

```{r mp, cache = TRUE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by Algorithm 2.', fig.dpi = 10}
gridExtra::grid.arrange(plot.A(A.bmp, zhat.bmp),
                        plot.A(A.pb, zhat.pb),
                        plot.A(A.dblp, zhat.dblp),
                        ncol = 3)
```

@@

```{r, cache = TRUE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2.', fig.height = 2}
A.hh <- readRDS('data/village-hh-59.rds')
zhat.hh <- cluster.pabm(A.hh, 6)
plot.A(A.hh, zhat.hh)
```

>>>>

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}

@@

*Proof*. 
Let $\Lambda^{(k)} = 
\begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
\in \mathbb{R}^{n_k \times K}$.  
Let $X$ be a block diagonal matrix 
$X = \blockdiag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}$.  
Let $L^{(k)}$ be a block diagonal matrix of column vectors $\lambda^{(lk)}$ for 
$l = 1, ..., K$. $L^{(k)} = \blockdiag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
\mathbb{R}^{n \times K}$.  
Let $Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
\mathbb{R}^{n \times K^2}$.  
Then $P = X Y^\top$.  
Similar to the $K = 2$ case, we again have $Y = X \Pi$ for a permutation matrix
$\Pi$, so $P = X \Pi X^\top$.  
The permutation described by $\Pi$ has $K$ fixed points, which correspond to 
$K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where 
$k = r (K + 1) + 1$ for $r = 0, ..., K - 1$. It also has 
$\binom{K}{2} = K (K - 1) / 2$ cycles of order $2$. Each cycle corresponds to 
a pair of eigenvalues $+1$ and $-1$ and a pair of eigenvectors 
$(e_s + e_t) / \sqrt{2}$ and $(e_s - e_t) / \sqrt{2}$.

@@

```{r mp, cache = FALSE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by Algorithm 2.', fig.dpi = 10}
gridExtra::grid.arrange(plot.A(A.bmp, zhat.bmp),
                        plot.A(A.pb, zhat.pb),
                        plot.A(A.dblp, zhat.dblp),
                        ncol = 3)
```

@@

```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2.', fig.height = 2}
hh <- readRDS('data/village-hh-59.rds')
A.hh <- hh$A
z.hh <- hh$z
K <- 6
zhat.hh <- cluster.pabm(A.hh, K, 
                        use.all = TRUE, normalize = FALSE, p = K, q = 0)

zhat.mres <- readRDS('data/mres-result.rds')

gridExtra::grid.arrange(plot.A(A.hh, zhat.hh), 
                        plot.A(A.hh, zhat.mres),
                        ncol = 2)
```
